---
title: "Capensis stranding model (one model only)"
output: html_notebook
---

Using the results in capensis_stranding_model.Rmd, I focus on one model in this document. It can be any model. When using v4b, the model uses Poisson distribution for the number of stranded long-beaked common dolphins along Silver Strand, whereas the negative binomial distribution is used for all beaches except Silver Strand in San Diego County. The additional mortality rate along SS is time dependent ($\mu_{1,t}$), whereas for other beaches it is constant ($\mu_2$).


```{r setup, include=F}
rm(list=ls())
library(jagsUI)
library(tidyverse)
library(ggplot2)
library(bayesplot)
library(loo)

source("SoCal_Strandings_fcns.R")
# MCMC setup
MCMC.params <- list(n.samples = 150000,
                    n.burnin = 125000,
                    n.thin = 5,
                    n.chains = 5)

n.per.chain <- (MCMC.params$n.samples - MCMC.params$n.burnin)/MCMC.params$n.thin

model.name <- "v4c"

# load data on capensis strandings
capensis.data <- read.csv(file = "data/capensis_strandings.csv")

# coast line lengths
SD.coast.length <- 150403.3 # meters from cetacean_v1.Rmd (SD.coast.no.MB.SDB.length)
SS.length <- 22135.74       # meters from cetacean_v1.Rmd (SS.length)
# from Becker et al. 2020. Habitat based density estimates for cetaceans in CA current ecosystem.
# 2001,53044
# 2005,52356
# 2008,58624
# 2014,58794
# 2018,83379
N <- c(NA, 53044, NA, NA, NA,
       52356, NA, NA, 58624,
       rep(NA, 5), 58794, rep(NA, 3), 83379, NA)

D1 <- as.vector(capensis.data$n.SS)
D1[is.na(D1)] <- 0

D2 <- as.vector(capensis.data$n.noSS)
D1_idx <- vector(mode = "numeric", length = length(D1))
D1_idx[D1>3] <- 1

# remove the first so we can start with known N
jags.data <- list(D1 = D1[2:length(D1)],
                  D2 = D2[2:length(D2)],
                  P_SS = SS.length/SD.coast.length,
                  T = length(D1) - 1,
                  N = N[2:length(N)],
                  D1_idx = D1_idx[2:length(N)])

parameters.list <- c("r", "f1", "f2", "mu", 
                     "mu1",  
                     "b", "sigma_N", "mean_N",
                     "mu_D1", "mu_D2", 
                     "Deviance", "loglik",
                     "p2")


jm <- run.model(parameters.list, jags.data, model.name, MCMC.params)


```



```{r}
best.model <- jm
best.model.MCMC <- jm$jm.MCMC

best.model.MCMC$Rhat
```

All parameters converged okay.

Take a look at Pareto k statistics.

```{r}
plot(best.model.MCMC$loo.out)

```

It looks okay. NOt great but okay. The SS portion is worse than the non-SS area.

```{r}
best.model.MCMC$loo.out
```


```{r}
mcmc_dens(best.model$jm.out$samples, c("r", "f", "mu",  "b", "mu1"))
```


```{r}
#mcmc_dens(best.model$samples, c("mu1", "p2"))
```



```{r}
# mcmc_dens(best.model$samples, 
#           c("mu1[1]", "mu1[2]", "mu1[3]", "mu1[4]",
#             "mu1[5]", "mu1[6]", "mu1[7]", "mu1[8]",
#             "mu1[9]"))
```

These posteriors don't look so good...


```{r}
# mcmc_trace(best.model$samples, 
#           c("mu1[1]", "mu1[2]", "mu1[3]", "mu1[4]",
#             "mu1[5]", "mu1[6]", "mu1[7]", "mu1[8]",
#             "mu1[9]"))
```

```{r}
# mcmc_dens(best.model$samples, 
#           c("mu1[10]", "mu1[11]", "mu1[12]", "mu1[13]",
#             "mu1[14]", "mu1[15]", "mu1[16]", "mu1[17]",
#             "mu1[18]", "mu1[19]"))
```

Some posteriors (10, 11,, 12, 15, and 19) seemed to be pushed against the upper limit (0.2). 

```{r}
# mcmc_trace(best.model$samples, 
#            c("mu1[10]", "mu1[11]", "mu1[12]", "mu1[13]",
#              "mu1[14]", "mu1[15]", "mu1[16]", "mu1[17]",
#              "mu1[18]", "mu1[19]"))
```



```{r}

mean.mu_D1 <- best.model$jm.out$mean$mu_D1
low.mu_D1 <- best.model$jm.out$q2.5$mu_D1
high.mu_D1 <- best.model$jm.out$q97.5$mu_D1

mean.mu_D2 <- best.model$jm.out$mean$mu_D2
low.mu_D2 <- best.model$jm.out$q2.5$mu_D2
high.mu_D2 <- best.model$jm.out$q97.5$mu_D2

mean.mu1 <- best.model$jm.out$mean$mu1
low.mu1 <- best.model$jm.out$q2.5$mu1
high.mu1 <- best.model$jm.out$q97.5$mu1

mean.mu2 <- best.model$jm.out$mean$mu2
low.mu2 <- best.model$jm.out$q2.5$mu2
high.mu2 <- best.model$jm.out$q97.5$mu2

ggplot(data.frame(D1 = jags.data$D1, 
                  mu_D1 = mean.mu_D1,
                  high.mu_D1 = high.mu_D1,
                  low.mu_D1 = low.mu_D1,
                  time = seq(2001, 2019))) +
  geom_point(aes(x = time, y = D1), color = "red") +
  geom_point(aes(x = time, y = mu_D1), color = "blue") +
  geom_errorbar(aes(x = time, ymin = low.mu_D1, ymax = high.mu_D1))

```

Red dots are the observed number of strandings along SS.


Observed # strandings vs. estimated mortality along SS. Upper limits are pushed against 0.2. 

```{r}
ggplot(data.frame(D1 = jags.data$D1, 
                  mu1 = mean.mu1, 
                  mu_D1 = mean.mu_D1, 
                  low_mu1 = low.mu1,
                  high_mu1 = high.mu1,
                  time = seq(2001, 2019))) +
  geom_point(aes(x = D1, y = mu1), color = "red") +
  geom_errorbar(aes(x = D1, ymin = low_mu1, ymax = high_mu1))


```

The following is not very useful as mu2 is constant over time. 

<!-- ```{r} -->
<!-- ggplot(data.frame(D2 = jags.data$D2,  -->
<!--                   mu2 = mean.mu2,  -->
<!--                   mu_D2 = mean.mu_D2,  -->
<!--                   low_mu2 = low.mu2, -->
<!--                   high_mu2 = high.mu2, -->
<!--                   time = seq(2001, 2019))) + -->
<!--   geom_point(aes(x = D2, y = mu2), color = "red") + -->
<!--   geom_errorbar(aes(x = D2, ymin = low_mu2, ymax = high_mu2)) -->

<!-- ``` -->


```{r}

ggplot(data.frame(D2 = jags.data$D2, 
                  mu_D2 = mean.mu_D2,
                  high.mu_D2 = high.mu_D2,
                  low.mu_D2 = low.mu_D2,
                  time = seq(2001, 2019))) +
  geom_point(aes(x = time, y = D2), color = "red") +
  geom_point(aes(x = time, y = mu_D2), color = "blue") +
  geom_errorbar(aes(x = time, ymin = low.mu_D2, ymax = high.mu_D2))

```

```{r}
mean.mean_N <- best.model$jm.out$mean$mean_N
low.mean_N <- best.model$jm.out$q2.5$mean_N
high.mean_N <- best.model$jm.out$q97.5$mean_N

ggplot(data.frame(N = jags.data$N, 
                  mu_N = mean.mean_N, 
                  low.mu_N = low.mean_N,
                  high.mu_N = high.mean_N,
                  time = seq(2001, 2019))) +
  geom_point(aes(x = time, y = mean.mean_N)) +
  geom_errorbar(aes(x = time, ymin = low.mean_N, ymax = high.mean_N)) + 
  geom_point(aes(x = time, y = N), color = "red", size = 2)

```
