---
title: "R Notebook"
output: html_notebook
---

This document an initial analysis of cetacean strandings along the coast of San Diego county. We are interested in looking at if there is any temporal and spatial patterns of strandings, especially related to military bases. 

In v2, I look at finer temporal scales - months, rather than years. 

```{r setup}
rm(list=ls())
library(tidyverse)
library(readr)
library(ggplot2)
library(ggridges)
library(sf)
#library(viridis)

source("SoCal_Strandings_fcns.R")
S.lat <- 32.4
N.lat <- 33.55
W.lon <- -118
E.lon <- -116.8

US.MX.border.lat <- 32.5344
TJriver.lat <- 32.55
SilverStrandN.lat <- 32.69
DogBeach.Nend.lat <- 32.685
Navy.base.S.lat <- 32.6369
Navy.base.N.lat <- 32.6742

col_defs <- cols(Specimen = col_character(),
                 RecoveryConditionID = col_character(),
                 IsSDCounty = col_character(),
                 SpeciesID = col_character(),
                 COMMONNAME = col_character(),
                 GENUS = col_factor(),
                 SPECIES = col_factor(),
                 OriginID = col_character(),
                 OriginDescription = col_character(),
                 SDCountyMapID = col_factor(),
                 Location = col_character(),
                 Year = col_integer(),
                 Month = col_integer(),
                 Day = col_integer(),
                 Latitude = col_double(),
                 Latitude_Precision = col_double(),
                 Latitude_Precision_Unit = col_character(),
                 Longitude = col_double(),
                 Longitude_Precision = col_double(),
                 Longitude_Precision_Unit = col_character())

# remove those with NA entries in lat/lon and San Diego Bay in Location
dat.1 <- read_csv(file = "data/SDCetaStrandings2000_2019.csv",
                  col_types = col_defs) %>% 
  filter(!is.na(Latitude)) %>%
  filter(Location != "San Diego Bay") %>%  # remove SDB
  filter(Longitude > -118)       # remove San Clemente stranding

# create a spatial object
dat.1 %>% select(Longitude, Latitude, Location,
                 Specimen, GENUS, SPECIES, COMMONNAME, 
                 SDCountyMapID, Year, Month) %>% 
  unite(col = "Year_Month", Year:Month, remove = F) %>%
  mutate(Decimal_Date = lubridate::decimal_date(lubridate::ym(Year_Month))) %>%
  st_as_sf(coords = c("Longitude", "Latitude"), 
           crs = "+proj=longlat +datum=WGS84") -> dat.1.sp

```

First look at the area map to orient ourselves. Some lat/lon seemed to be a bit off so went through one year at a time to see which ones needed to be corrected. Lines that are commented out were for that purpose. 

Bring in shapefiles:

```{r}
# get the coast lines
# Does this work with Windows?
home.dir <- Sys.getenv("HOMEPATH")
coast.line <- st_read(paste0(home.dir, "/Documents/Oceans and Maps/San Diego/COASTLINE"),
                      layer = "coast") %>% 
  st_transform(crs = "+proj=longlat +datum=WGS84") 

# this is for plotting:
harbor_bay_1 <- st_read(paste0(home.dir, "/Documents/Oceans and Maps/San Diego/Harbor_Bay"), 
                        layer = "Harbor_Bay") %>% 
   st_transform(crs = "+proj=longlat +datum=WGS84") 

# I edited SDB in QGIS on my laptop (linux) so that Silver Strand is not included.
# this is for computing coast length
harbor_bay <- st_read(paste0(home.dir, "/Documents/Oceans and Maps/San Diego/SDB_v2"),
               layer = "SDB_v2") %>% 
  st_transform(crs = "+proj=longlat +datum=WGS84") 

Pacific <-  st_read(paste0(home.dir, "/Documents/Oceans and Maps/San Diego/Pacific_Ocean"), 
                    layer = "PACIFIC_OCEAN") %>% 
  st_transform(crs = "+proj=longlat +datum=WGS84") 

# get county lines
county.lines <- st_read("data/ca-county-boundaries/CA_Counties")

county.lines %>% filter(NAME == "San Diego") %>%
  st_transform(crs = "+proj=longlat +datum=WGS84") -> SD.county.line

# Silver Strand - created from the entire coast line above in QGIS
SS <- st_read(paste0(home.dir, "/Documents/Oceans and Maps/San Diego/SilverStrand"),
               layer = "SilverStrand") %>% 
  st_transform(crs = "+proj=longlat +datum=WGS84") %>%
  st_cast("LINESTRING")

Navy <- st_read(paste0(home.dir, "/Documents/Oceans and Maps/Navy_installations_poly"), 
                layer = "Navy_installations_poly") %>% 
  st_transform(crs = "+proj=longlat +datum=WGS84") %>%
  st_make_valid()  # some duplicated vertices removed. 

# https://gis.stackexchange.com/questions/404385/r-sf-some-edges-are-crossing-in-a-multipolygon-how-to-make-it-valid-when-using/404454


land.color <- "cornsilk" #'#333333'
water.color <- "lightcyan1"
Navy.color <- "salmon"
font.size <- 3.5

```



Exclude those that are found outside of SD county

```{r}
in.SD <- st_join(dat.1.sp, SD.county.line) %>% 
   filter(!is.na(AWATER)) 

ggplot() +
  geom_sf(data = Pacific, fill = water.color, color = water.color) +
  geom_sf(data = harbor_bay_1, fill = water.color) +
  geom_sf(data = coast.line) +
  
  geom_sf(data = SD.county.line,
          color = "purple",
          fill = "transparent") +
  geom_sf(data = Navy, fill = Navy.color) +

  geom_sf(data = in.SD, 
          aes(color = as.factor(Year))) +
  coord_sf(xlim = c(W.lon, E.lon),
           ylim = c(S.lat, N.lat)) +
  geom_text(data = data.frame(x = -117.1, y = 33.3),
            aes(x = x, y = y, label = "San Diego"),
            color = "black", 
            size = font.size, fontface = "bold") + # angle = -45 
  ylab("Latitude") +
  xlab("Longitude") +
  ggtitle("Cetacean strandings along San Diego County (2000-2019)") +
  #scale_fill_continuous(name = "Map ID") +
  theme(plot.title = element_text(hjust = 0.5),
        #legend.position = "top", #c(0.1, 0.3),
        legend.background = element_blank(),
        legend.title = element_text("Map ID"),
        panel.background = element_rect(fill = land.color))

# ggsave(filename = "figures/StrandingLocations.png",
#        device = "png",
#        dpi = 600)


```


Plot strandings as a function of month. 

```{r}
#Y <- seq(2000, 2019)
#Y <- 2019

ggplot() +
  geom_sf(data = Pacific, fill = water.color, color = water.color) +
  geom_sf(data = harbor_bay_1, fill = water.color) +
  geom_sf(data = coast.line) +
  
  geom_sf(data = SD.county.line,
          color = "purple",
          fill = "transparent") +
    geom_sf(data = Navy, fill = Navy.color) +

  geom_sf(data = in.SD, 
          aes(color = as.factor(Month)), alpha = 0.6) +
  coord_sf(xlim = c(W.lon, E.lon),
           ylim = c(S.lat, N.lat)) +
  geom_text(data = data.frame(x = -117.1, y = 33.3),
            aes(x = x, y = y, label = "San Diego"),
            color = "black", 
            size = font.size, fontface = "bold") + 
  ylab("Latitude") +
  xlab("Longitude") +
  ggtitle("Cetacean strandings along San Diego County (2000-2019)") +
  theme(plot.title = element_text(hjust = 0.5),
        #legend.position = "top", #c(0.1, 0.3),
        legend.background = element_blank(),
        legend.title = element_blank(),
        panel.background = element_rect(fill = land.color))


```


Just Silver Strand:

```{r}
W.lon.SS <- -117.35
E.lon.SS <- -117.05
SDB.mouth.lon <- -117.23
N.lat.SS <- 32.73
S.lat.SS <- US.MX.border.lat

dat.1 %>% filter(Latitude < SilverStrandN.lat & 
                   Latitude > US.MX.border.lat &
                   Longitude > SDB.mouth.lon) %>% 
  st_as_sf(coords = c("Longitude", "Latitude"), 
           crs = "+proj=longlat +datum=WGS84") -> dat.1.SS.sp

ggplot() +
  geom_sf(data = coast.line)+
  geom_sf(data = harbor_bay_1, fill = water.color) +
  geom_sf(data = Pacific, fill = water.color) +
  geom_sf(data = SD.county.line,
          color = "purple",
          fill = "transparent") +
  
  geom_sf(data = Navy, fill = Navy.color) +
  
  geom_sf(data = dat.1.SS.sp, 
          aes(color = as.factor(Month)), alpha = 0.6) +
  #geom_sf(data = in.bay.2, color = "black", size = 1.2) +
  coord_sf(xlim = c(W.lon.SS, -117.05),
           ylim = c(S.lat.SS, N.lat.SS)) +
  ylab("Latitude") +
  xlab("Longitude") +
  ggtitle("Cetacean strandings along Silver Strand (2000-2019)") +
  theme(plot.title = element_text(hjust = 0.5),
        #legend.position = "top", #c(0.1, 0.3),
        legend.background = element_blank(),
        legend.title = element_blank(),
        panel.background = element_rect(fill = land.color))

# ggsave(p.2, filename = "figures/SS_strandings.png",
#        device = "png", dpi = 600)
```


Simple statistics of # strandings per unit coast length

```{r}
# first, intersection between coast line and county
SD.coast <- st_intersection(coast.line, SD.county.line)

# need to remove Mission Bay and San Diego Bay
# MB - convert polygon to line
harbor_bay %>% filter(NAME == "MISSION BAY") -> MB.poly
MB.line <- st_cast(MB.poly, "LINESTRING")
MB.length <- sum(st_length(MB.line))

# SDB coast length
harbor_bay %>% filter(NAME == "SAN DIEGO BAY") -> SDB.poly
SDB.line <- st_cast(SDB.poly, "LINESTRING")
SDB.length <- sum(st_length(SDB.line))

SD.coast.no.MB.SDB.length <- as.numeric(sum(st_length(SD.coast)) - MB.length - SDB.length)

# Silver Strand coast length
SS.length <- as.numeric(st_length(SS))

# remove offshore findings
in.SD.sp <- st_join(dat.1.sp, SD.county.line) %>% 
   filter(!is.na(AWATER)) 

in.SD.sp %>% as.data.frame() %>%
  unite(col = "Sp", GENUS:SPECIES) %>%
  group_by(Sp) %>%
  summarise(n = n()) %>% 
  arrange(desc(n)) -> in.SD.bySp

in.SD.sp %>% filter(SPECIES == "capensis") -> in.SD.capensis

# remove Silver Strand strandings
in.SD.capensis %>% filter(Location != "Imperial Beach/Coronado") -> in.SD.capensis.noSS

in.SD.capensis %>% filter(Location == "Imperial Beach/Coronado") -> in.SD.capensis.SS

in.SD.capensis %>% as.data.frame() %>%
  select(Year, Month) %>%
  group_by(Year, Month) %>%
  summarise(n = n(),
            n_per_km = n/((SD.coast.no.MB.SDB.length)/1000),
            Coast = "All") %>%
  unite(col = "Year_Month", Year:Month, remove = F) %>%
  mutate(Decimal_Date = lubridate::decimal_date(lubridate::ym(Year_Month))) -> capensis.perKm.all


in.SD.capensis.noSS %>% as.data.frame() %>%
  select(Year, Month) %>%
  group_by(Year, Month) %>%
  summarise(n = n(),
            n_per_km = n/((SD.coast.no.MB.SDB.length-SS.length)/1000),
            Coast = "No SS") %>%
  unite(col = "Year_Month", Year:Month, remove = F) %>%
  mutate(Decimal_Date = lubridate::decimal_date(lubridate::ym(Year_Month))) -> capensis.perKm.noSS

in.SD.capensis.SS %>% as.data.frame() %>%
  select(Year, Month) %>%
  group_by(Year, Month) %>%
  summarise(n = n(),
            n_per_km = n/((SS.length)/1000),
            Coast = "SS only") %>%
  unite(col = "Year_Month", Year:Month, remove = F) %>%
  mutate(Decimal_Date = lubridate::decimal_date(lubridate::ym(Year_Month))) -> capensis.perKm.SS

capensis.perKm <- rbind(capensis.perKm.all, 
                         capensis.perKm.noSS, 
                         capensis.perKm.SS)

capensis.perKm.all %>% 
  select(Year, n, n_per_km) %>% 
  transmute("Year" = Year, "n.all" = n, "n_per_km_all" = n_per_km) %>% #-> tmp
  left_join(capensis.perKm.noSS, by = "Year") %>%
  transmute("Year" = Year, "n.all" = n.all, "n_per_km_all" = n_per_km_all, 
            "n.noSS" = n, "n_per_km_noSS" = n_per_km) %>%
  left_join(capensis.perKm.SS, by = "Year") %>%
  transmute("Year" = Year, "n.all" = n.all, "n_per_km_all" = n_per_km_all,
            "n.noSS" = n.noSS, "n_per_km_noSS" = n_per_km_noSS, 
            "n.SS" = n, "n_per_km_SS" = n_per_km) -> capensis.perKm.wide


ggplot(capensis.perKm) +
  geom_bar(aes(x = Year, 
               y = n_per_km, 
               fill = Coast),
           stat = "identity",
           position = position_dodge())  + 
  scale_fill_viridis_d() +
  theme(legend.position = "top",
        legend.title = element_blank(),
        axis.title.x = element_blank()) +
  ylab("# strandings per km of coast line")

# ggsave(filename = "figures/strandingPerKm.png",
#        device = "png", dpi = 600)
```


Look at what affects the stranding rate:

NOI: Northern and Southern Oscillation Indices - The NOI (extratropical-based Northern Oscillation Index) and its analog, the SOI* (extratropical-based Southern Oscillation Index) are new indices of midlatitude climate fluctuations that show interesting relationships with fluctuations in marine ecosystems and populations. They reflect the variability in equatorial and extratropical teleconnections and represent a wide range of local and remote climate signals. The indices are counterparts to the traditional SOI (Southern Oscillation Index) that relate variability in the atmospheric forcing of climate change in northern and southern midlatitude hemisphere regions. (https://oceanview.pfeg.noaa.gov/erddap/info/cciea_OC_NOI/index.html)

PDO: Updated standardized values for the PDO index, derived as the leading PC of monthly SST anomalies in the North Pacific Ocean, poleward of 20N. The monthly mean global average SST anomalies are removed to separate this pattern of variability from any "global warming" signal that may be present in the data. (https://oceanview.pfeg.noaa.gov/erddap/info/cciea_OC_PDO/index.html)

```{r}

NOI <- read_csv(file = "data/NOI_v1.csv",
                col_types = cols(time = col_datetime(),
                  NOI = col_double(),
                  SOIX = col_double(),
                  SOI = col_double())) %>%
  mutate(Year = lubridate::year(time),
         Month = lubridate::month(time)) %>%
  unite("Year_Month", Year:Month, remove = FALSE) %>%
  mutate(Decimal_Date = lubridate::decimal_date(lubridate::ym(Year_Month)))


PDO <- read_csv(file = "data/PDO_v1.csv",
                col_types = cols(time = col_datetime())) %>%
  mutate(Year = lubridate::year(time),
         Month = lubridate::month(time)) %>%
  unite("Year_Month", Year:Month, remove = FALSE) %>%
  mutate(Decimal_Date = lubridate::decimal_date(lubridate::ym(Year_Month)))


capensis.perKm.all %>% left_join(NOI, by = "Year_Month") %>%
  left_join(PDO, by = "Year_Month") -> capensis.perKm.all.NOI.PDO

ggplot(capensis.perKm.all.NOI.PDO) +
  geom_path(aes(x = Decimal_Date, y = NOI)) +
  geom_point(aes(x = Decimal_Date, y = NOI, color = n_per_km)) +
  scale_color_viridis_c()
  


```


```{r}
ggplot(capensis.perKm.all.NOI.PDO) +
  geom_point(aes(y = n_per_km, x = NOI))
```



```{r}
ggplot(capensis.perKm.all.NOI.PDO) +
  geom_path(aes(x = Decimal_Date, y = PDO)) +
  geom_point(aes(x = Decimal_Date, y = PDO, color = n_per_km)) +
  scale_color_viridis_c()
  
```

```{r}


ggplot(capensis.perKm.all.NOI.PDO) +
  geom_point(aes(y = n_per_km, x = PDO))

```


```{r}
capensis.perKm.SS %>% left_join(NOI, by = "Year_Month") %>%
  left_join(PDO, by = "Year_Month") -> capensis.perKm.SS.NOI.PDO

ggplot(capensis.perKm.SS.NOI.PDO) +
 geom_path(aes(x = Decimal_Date, y = NOI)) +
  geom_point(aes(x = Decimal_Date, y = NOI, color = n_per_km)) +
  scale_color_viridis_c()
  
```


```{r}
ggplot(capensis.perKm.SS.NOI.PDO) +
  geom_point(aes(y = n_per_km, x = NOI))
```


```{r}
ggplot(capensis.perKm.SS.NOI.PDO) +
 geom_path(aes(x = Decimal_Date, y = PDO)) +
  geom_point(aes(x = Decimal_Date, y = PDO, color = n_per_km)) +
  scale_color_viridis_c()

```


```{r}
ggplot(capensis.perKm.SS.NOI.PDO) +
  geom_path(aes(x = Decimal_Date, y = PDO)) +
  geom_point(aes(x = Decimal_Date, y = PDO, color = n_per_km)) +
  scale_color_viridis_c()
  
```


```{r}

ggplot(capensis.perKm.SS.NOI.PDO) +
  geom_point(aes(y = n_per_km, x = PDO))

```

There is no apparent relationship between #/km and NOI/PDO. I guess it can't be that simple...

