---
title: "R Notebook"
output: html_notebook
---

This document an initial analysis of cetacean strandings along the coast of San Diego county. We are interested in looking at if there is any temporal and spatial patterns of strandings, especially related to military bases. 

```{r setup}
rm(list=ls())
library(tidyverse)
library(readr)
library(ggplot2)
library(ggridges)
library(sf)
#library(viridis)

source("SoCal_Strandings_fcns.R")
S.lat <- 32.4
N.lat <- 33.55
W.lon <- -118
E.lon <- -116.8

US.MX.border.lat <- 32.5344
TJriver.lat <- 32.55
SilverStrandN.lat <- 32.69
DogBeach.Nend.lat <- 32.685
Navy.base.S.lat <- 32.6369
Navy.base.N.lat <- 32.6742

col_defs <- cols(Specimen = col_character(),
                 RecoveryConditionID = col_character(),
                 IsSDCounty = col_character(),
                 SpeciesID = col_character(),
                 COMMONNAME = col_character(),
                 GENUS = col_factor(),
                 SPECIES = col_factor(),
                 OriginID = col_character(),
                 OriginDescription = col_character(),
                 SDCountyMapID = col_factor(),
                 Location = col_character(),
                 Year = col_integer(),
                 Month = col_integer(),
                 Day = col_integer(),
                 Latitude = col_double(),
                 Latitude_Precision = col_double(),
                 Latitude_Precision_Unit = col_character(),
                 Longitude = col_double(),
                 Longitude_Precision = col_double(),
                 Longitude_Precision_Unit = col_character())

# remove those with NA entries in lat/lon and San Diego Bay in Location
dat.1 <- read_csv(file = "data/SDCetaStrandings2000_2019.csv",
                  col_types = col_defs) %>% 
  filter(!is.na(Latitude)) %>%
  filter(Location != "San Diego Bay") %>%
  filter(Longitude > -118)       # remove San Clemente stranding

# create a spatial object
dat.1 %>% select(Longitude, Latitude, 
                 Specimen, GENUS, SPECIES, COMMONNAME, 
                 SDCountyMapID, Year, Month) %>% 

  st_as_sf(coords = c("Longitude", "Latitude"), 
           crs = "+proj=longlat +datum=WGS84") -> dat.1.sp

```

First look at the area map to orient ourselves. Some lat/lon seemed to be a bit off so went through one year at a time to see which ones needed to be corrected. Lines that are commented out were for that purpose. 

Bring in shapefiles:

```{r}
# get the coast lines
# Does this work with Windows?
home.dir <- Sys.getenv("HOMEPATH")
coast.line <- st_read(paste0(home.dir, "/Documents/Oceans and Maps/San Diego/COASTLINE"),
                      layer = "coast") %>% 
  st_transform(crs = "+proj=longlat +datum=WGS84") 

# this is for plotting:
harbor_bay_1 <- st_read(paste0(home.dir, "/Documents/Oceans and Maps/San Diego/Harbor_Bay"), 
                        layer = "Harbor_Bay") %>% 
   st_transform(crs = "+proj=longlat +datum=WGS84") 

# I edited SDB in QGIS on my laptop (linux) so that Silver Strand is not included.
# this is for computing coast length
harbor_bay <- st_read(paste0(home.dir, "/Documents/Oceans and Maps/San Diego/SDB_v2"),
               layer = "SDB_v2") %>% 
  st_transform(crs = "+proj=longlat +datum=WGS84") 

Pacific <-  st_read(paste0(home.dir, "/Documents/Oceans and Maps/San Diego/Pacific_Ocean"), 
                    layer = "PACIFIC_OCEAN") %>% 
  st_transform(crs = "+proj=longlat +datum=WGS84") 

# get county lines
county.lines <- st_read("data/ca-county-boundaries/CA_Counties")

county.lines %>% filter(NAME == "San Diego") %>%
  st_transform(crs = "+proj=longlat +datum=WGS84") -> SD.county.line

# Silver Strand - created from the entire coast line above in QGIS
SS <- st_read(paste0(home.dir, "/Documents/Oceans and Maps/San Diego/SilverStrand"),
               layer = "SilverStrand") %>% 
  st_transform(crs = "+proj=longlat +datum=WGS84") %>%
  st_cast("LINESTRING")

Navy <- st_read(paste0(home.dir, "/Documents/Oceans and Maps/Navy_installations_poly"), 
                layer = "Navy_installations_poly") %>% 
  st_transform(crs = "+proj=longlat +datum=WGS84") %>%
  st_make_valid()  # some duplicated vertices removed. 

# https://gis.stackexchange.com/questions/404385/r-sf-some-edges-are-crossing-in-a-multipolygon-how-to-make-it-valid-when-using/404454


land.color <- "cornsilk" #'#333333'
water.color <- "lightcyan1"
Navy.color <- "salmon"
font.size <- 3.5

```


Put maps together:

```{r}
ggplot() +
  geom_sf(data = Pacific, fill = water.color, color = water.color) +
  geom_sf(data = harbor_bay_1, fill = water.color) +
  geom_sf(data = coast.line) +
  
  geom_sf(data = SD.county.line,
          color = "purple",
          fill = "transparent") +
  geom_sf(data = Navy, fill = Navy.color) +
  geom_text(data = data.frame(x = -117.1, y = 33.3),
            aes(x = x, y = y, label = "San Diego"),
            color = "black", 
            size = font.size, fontface = "bold") + # angle = -45 

  #geom_sf(data = dat.1.sp, 
  #        aes(color = SDCountyMapID)) +
  coord_sf(xlim = c(W.lon, E.lon),
           ylim = c(S.lat, N.lat)) +
  ylab("Latitude") +
  xlab("Longitude") +
  theme(#plot.title = element_text(hjust = 0.5),
        #legend.position = "top", #c(0.1, 0.3),
        #legend.background = element_blank(),
        #legend.title = element_text("Map ID"),
        panel.background = element_rect(fill = land.color))

#ggsave(filename = "figures/SDcoast.png", device = "png", dpi = 600)

```


Figuring out Map IDs:

```{r}
ggplot() +
  geom_sf(data = Pacific, fill = water.color, color = water.color) +
  geom_sf(data = harbor_bay_1, fill = water.color) +
  geom_sf(data = coast.line) +
  
  geom_sf(data = SD.county.line,
          color = "purple",
          fill = "transparent") +
  geom_sf(data = Navy, fill = Navy.color) +

  geom_sf(data = dat.1.sp, 
          aes(color = SDCountyMapID)) +
  coord_sf(xlim = c(W.lon, E.lon),
           ylim = c(S.lat, N.lat)) +
  geom_text(data = data.frame(x = -117.1, y = 33.3),
            aes(x = x, y = y, label = "San Diego"),
            color = "black", 
            size = font.size, fontface = "bold") + # angle = -45 
  ylab("Latitude") +
  xlab("Longitude") +
  ggtitle("Cetacean strandings along San Diego County (2000-2019)") +
  #scale_fill_continuous(name = "Map ID") +
  theme(plot.title = element_text(hjust = 0.5),
        #legend.position = "top", #c(0.1, 0.3),
        legend.background = element_blank(),
        legend.title = element_text("Map ID"),
        panel.background = element_rect(fill = land.color))


```

San Diego Bay should be area 2. There are still some inconsistent data... Kerri is looking into it. 2021-09-07 

I think these are all fixed. 2021-09-09

Exclude those that are found outside of SD county

```{r}
in.SD <- st_join(dat.1.sp, SD.county.line) %>% 
   filter(!is.na(AWATER)) 

ggplot() +
  geom_sf(data = Pacific, fill = water.color, color = water.color) +
  geom_sf(data = harbor_bay_1, fill = water.color) +
  geom_sf(data = coast.line) +
  
  geom_sf(data = SD.county.line,
          color = "purple",
          fill = "transparent") +
  geom_sf(data = Navy, fill = Navy.color) +

  geom_sf(data = in.SD, 
          aes(color = as.factor(Year))) +
  coord_sf(xlim = c(W.lon, E.lon),
           ylim = c(S.lat, N.lat)) +
  geom_text(data = data.frame(x = -117.1, y = 33.3),
            aes(x = x, y = y, label = "San Diego"),
            color = "black", 
            size = font.size, fontface = "bold") + # angle = -45 
  ylab("Latitude") +
  xlab("Longitude") +
  ggtitle("Cetacean strandings along San Diego County (2000-2019)") +
  #scale_fill_continuous(name = "Map ID") +
  theme(plot.title = element_text(hjust = 0.5),
        #legend.position = "top", #c(0.1, 0.3),
        legend.background = element_blank(),
        legend.title = element_text("Map ID"),
        panel.background = element_rect(fill = land.color))

# ggsave(filename = "figures/StrandingLocations.png",
#        device = "png",
#        dpi = 600)


```


Plot strandings as a function of year. 

```{r}
#Y <- seq(2000, 2019)
#Y <- 2019
k <- 1
#for (k in 1:length(Y)){
ggplot() +
  geom_sf(data = Pacific, fill = water.color, color = water.color) +
  geom_sf(data = harbor_bay_1, fill = water.color) +
  geom_sf(data = coast.line) +
  
  geom_sf(data = SD.county.line,
          color = "purple",
          fill = "transparent") +
    geom_sf(data = Navy, fill = Navy.color) +

  geom_sf(data = dat.1.sp, 
          aes(color = as.factor(Year)), alpha = 0.6) +
  coord_sf(xlim = c(W.lon, E.lon),
           ylim = c(S.lat, N.lat)) +
  geom_text(data = data.frame(x = -117.1, y = 33.3),
            aes(x = x, y = y, label = "San Diego"),
            color = "black", 
            size = font.size, fontface = "bold") + # angle = -45 
  # geom_point(data = dat.1,
  #            aes(x = Longitude,
  #                y = Latitude,
  #                color = as.factor(Year)),
  #            alpha = 0.7,
  #            size = 3) +
  # geom_point(data = filter(dat.1, Year == Y),
  #            aes(x = Longitude, 
  #                y = Latitude),
  #            alpha = 0.7,
  #            size = 3) +
  ylab("Latitude") +
  xlab("Longitude") +
  ggtitle("Cetacean strandings along San Diego County (2000-2019)") +
  #ggtitle(paste0("Cetacean strandings ", Y[k])) +
  #theme(panel.grid = element_line(size = 1, color = "black"),
  #      panel.grid.minor = element_line(size = 0.5, color = "black")) + 
  #scale_y_continuous(minor_breaks = seq(S.lat, N.lat, 0.1))
  theme(plot.title = element_text(hjust = 0.5),
        #legend.position = "top", #c(0.1, 0.3),
        legend.background = element_blank(),
        legend.title = element_blank(),
        panel.background = element_rect(fill = land.color))

#ggsave(p.1, filename = paste0("figures/StrandingLocations_", Y[k], ".png"),
#       device = "png", dpi = 600)  
#}

# ggsave(filename = "figures/StrandingAllLocations.png",
#        device = "png",
#        dpi = 600)


```



Just Silver Strand:

```{r}
W.lon.SS <- -117.35
E.lon.SS <- -117.05
SDB.mouth.lon <- -117.23
N.lat.SS <- 32.73
S.lat.SS <- US.MX.border.lat

dat.1 %>% filter(Latitude < SilverStrandN.lat & 
                   Latitude > US.MX.border.lat &
                   Longitude > SDB.mouth.lon) %>% 
  st_as_sf(coords = c("Longitude", "Latitude"), 
           crs = "+proj=longlat +datum=WGS84") -> dat.1.SS.sp

# To find points that are in the bay, do the following:
# in.bay <- st_join(dat.1.SS.sp, harbor_bay) %>% 
#   filter(!is.na(SHAPE_AREA)) 

# San Diego Bay is defined to include the northern tip of Silver Strand
# along the beach of North Island Air Station. So, those need to be removed. 
# These have been fixed.
# in.bay %>%  
#   st_coordinates() %>%
#   data.frame() %>%
#   filter(Y < 32.67) %>%
#   row.names() -> in.bay.idx
# 
# in.bay.2 <- in.bay[as.numeric(in.bay.idx), ]


p.2 <- ggplot() +
  geom_sf(data = coast.line)+
  geom_sf(data = harbor_bay_1, fill = water.color) +
  geom_sf(data = Pacific, fill = water.color) +
  geom_sf(data = SD.county.line,
          color = "purple",
          fill = "transparent") +
  
  geom_sf(data = Navy, fill = Navy.color) +
  
  geom_sf(data = dat.1.SS.sp, 
          aes(color = as.factor(Year)), alpha = 0.6) +
  #geom_sf(data = in.bay.2, color = "black", size = 1.2) +
  coord_sf(xlim = c(W.lon.SS, -117.05),
           ylim = c(S.lat.SS, N.lat.SS)) +

  # geom_point(data = dat.1.SS,
  #            aes(x = Longitude,
  #                y = Latitude,
  #                color = as.factor(Year)),
  #            alpha = 0.7,
  #            size = 3) +
  # geom_point(data = filter(dat.1, Year == Y),
  #            aes(x = Longitude, 
  #                y = Latitude),
  #            alpha = 0.7,
  #            size = 3) +
  ylab("Latitude") +
  xlab("Longitude") +
  ggtitle("Cetacean strandings along Silver Strand (2000-2019)") +
  #ggtitle(paste0("Cetacean strandings ", Y[k])) +
  #theme(panel.grid = element_line(size = 1, color = "black"),
  #      panel.grid.minor = element_line(size = 0.5, color = "black")) + 
  #scale_y_continuous(minor_breaks = seq(S.lat, N.lat, 0.1))
  theme(plot.title = element_text(hjust = 0.5),
        #legend.position = "top", #c(0.1, 0.3),
        legend.background = element_blank(),
        legend.title = element_blank(),
        panel.background = element_rect(fill = land.color))

# ggsave(p.2, filename = "figures/SS_strandings.png",
#        device = "png", dpi = 600)
```


A quick plot of all strandings vs latitude:
```{r}

ggplot(data = dat.1) +
  geom_rect(aes(xmin = US.MX.border.lat, xmax = SilverStrandN.lat,
                ymin = 0, ymax = 10), alpha = 0.3, fill = "orange") +
  geom_histogram(aes(x = Latitude), binwidth = 0.03) +
  facet_wrap("Month") +
  labs(title = "Cetacean strandings in San Diego county by month (2000-2019)")

```

Simple statistics of # strandings per unit coast length

```{r}
# first, intersection between coast line and county
SD.coast <- st_intersection(coast.line, SD.county.line)

# need to remove Mission Bay and San Diego Bay
# MB - convert polygon to line
harbor_bay %>% filter(NAME == "MISSION BAY") -> MB.poly
MB.line <- st_cast(MB.poly, "LINESTRING")
MB.length <- sum(st_length(MB.line))

# SDB coast length
harbor_bay %>% filter(NAME == "SAN DIEGO BAY") -> SDB.poly
SDB.line <- st_cast(SDB.poly, "LINESTRING")
SDB.length <- sum(st_length(SDB.line))

SD.coast.no.MB.SDB.length <- as.numeric(sum(st_length(SD.coast)) - MB.length - SDB.length)

# Silver Strand coast length
SS.length <- as.numeric(st_length(SS))

# remove Silver Strand strandings
dat.1 %>% filter(Location != "Imperial Beach/Coronado") -> dat.1.noSS
dat.1.noSS %>% st_as_sf(coords = c("Longitude", "Latitude"), 
           crs = "+proj=longlat +datum=WGS84") -> dat.1.noSS.sp 

dat.1 %>% filter(Location == "Imperial Beach/Coronado") -> dat.1.SS
dat.1.SS %>% st_as_sf(coords = c("Longitude", "Latitude"), 
           crs = "+proj=longlat +datum=WGS84") -> dat.1.SS.sp 

dat.1 %>% group_by(Year) %>%
  summarise(n = n(),
            n_per_km = n/((SD.coast.no.MB.SDB.length)/1000),
            Coast = "All") -> stranding.perKm.all


dat.1.noSS %>% group_by(Year) %>%
  summarise(n = n(),
            n_per_km = n/((SD.coast.no.MB.SDB.length-SS.length)/1000),
            Coast = "No SS") -> stranding.perKm.noSS

dat.1.SS %>% group_by(Year) %>%
  summarise(n = n(),
            n_per_km = n/((SS.length)/1000),
            Coast = "SS only") -> stranding.perKm.SS

stranding.perKm <- rbind(stranding.perKm.all, 
                         stranding.perKm.noSS, 
                         stranding.perKm.SS)

ggplot(stranding.perKm) +
  geom_bar(aes(x = Year, 
               y = n_per_km, 
               fill = Coast),
           stat = "identity",
           position = position_dodge())  + 
  scale_fill_viridis_d() +
  theme(legend.position = "top",
        legend.title = element_blank(),
        axis.title.x = element_blank()) +
  ylab("# strandings per km of coast line")

# ggsave(filename = "figures/strandingPerKm.png",
#        device = "png", dpi = 600)
```


Split the dataset in two: one for 2000-2009 and another for 2010-2019.

```{r}
dat.1.1 <- filter(dat.1, Year < 2010)

ggplot(data = dat.1.1) +
  geom_rect(aes(xmin = TJriver.lat, xmax = SilverStrandN.lat,
                ymin = 0, ymax = 10), 
            alpha = 0.3, fill = "orange") +
  geom_histogram(aes(x = Latitude), binwidth = 0.03) +
  facet_wrap("Month") +
  labs(title = "Cetacean strandings in San Diego county by month (2000-2009)")


```


```{r}
dat.1.2 <- filter(dat.1, Year > 2009)

ggplot(data = dat.1.2) +
  geom_rect(aes(xmin = TJriver.lat, xmax = SilverStrandN.lat,
                ymin = 0, ymax = 10), 
            alpha = 0.3, fill = "orange") +
  geom_histogram(aes(x = Latitude), binwidth = 0.03) +
  facet_wrap("Month") +
  labs(title = "Cetacean strandings in San Diego county by month (2010-2019)")


```

As Nick was showing... Look at just Silver Strand:

```{r}

ggplot(data = dat.1.SS) +
  geom_rect(aes(xmin = US.MX.border.lat, xmax = TJriver.lat, 
                ymin = 0, ymax = 10), 
            alpha = 0.3, fill = "green") +
  geom_rect(aes(xmin = Navy.base.S.lat, xmax = Navy.base.N.lat,
                ymin = 0, ymax = 10), 
            alpha = 0.3, fill = "orange") +
  geom_rect(aes(xmin = DogBeach.Nend.lat, xmax = SilverStrandN.lat,
                ymin = 0, ymax = 10), 
            alpha = 0.3, fill = "purple") +
  geom_histogram(aes(x = Latitude), binwidth = 0.01) +
  facet_wrap("Month") +
  labs(title = "Cetacean strandings along Silver Strand by month (2000-2009)")



```

It is quite obvious that Silver Strand has higher rate of strandings than other parts of the SD county. 

Look at the species

```{r}
ggplot() +
  geom_sf(data = coast.line)+
  geom_sf(data = harbor_bay_1, fill = water.color) +
  geom_sf(data = Pacific, fill = water.color) +
    geom_sf(data = SD.county.line,
          color = "purple",
          fill = "transparent") +

  geom_sf(data = Navy, fill = Navy.color) +

  geom_sf(data = dat.1.sp, 
          aes(color = GENUS), alpha = 0.6) +
  coord_sf(xlim = c(W.lon, E.lon),
           ylim = c(S.lat, N.lat)) +

  ylab("Latitude") +
  xlab("Longitude") +
  ggtitle("Cetacean strandings along San Diego (2000-2019)") +

  # theme(panel.grid = element_line(size = 1, color = "black"),
  #       panel.grid.minor = element_line(size = 0.5, color = "black")) + 
  #scale_y_continuous(minor_breaks = seq(S.lat, N.lat, 0.1))
  theme(plot.title = element_text(hjust = 0.5),
        #legend.position = "top", #c(0.1, 0.3),
        legend.background = element_blank(),
        legend.title = element_blank(),
        panel.background = element_rect(fill = land.color))

```

Look at the number of strandings in and out of military installations

```{r}
# To find points that are in the military installations:

on.base <- st_join(dat.1.sp, Navy) %>%
  filter(!is.na(Shape_Area)) 

off.base <- st_join(dat.1.sp, Navy) %>%
  filter(is.na(Shape_Area))

ggplot() +
  geom_sf(data = coast.line)+
  geom_sf(data = harbor_bay_1, fill = water.color) +
  geom_sf(data = Pacific, fill = water.color) +
    geom_sf(data = SD.county.line,
          color = "purple",
          fill = "transparent") +

  geom_sf(data = Navy, fill = Navy.color) +

  geom_sf(data = on.base, 
          aes(color = GENUS), alpha = 0.6) +
  coord_sf(xlim = c(W.lon, E.lon),
           ylim = c(S.lat, N.lat)) +

  ylab("Latitude") +
  xlab("Longitude") +
  ggtitle("Cetacean strandings on military bases in San Diego (2000-2019)") +

  # theme(panel.grid = element_line(size = 1, color = "black"),
  #       panel.grid.minor = element_line(size = 0.5, color = "black")) + 
  #scale_y_continuous(minor_breaks = seq(S.lat, N.lat, 0.1))
  theme(plot.title = element_text(hjust = 0.5),
        #legend.position = "top", #c(0.1, 0.3),
        legend.background = element_blank(),
        legend.title = element_blank(),
        panel.background = element_rect(fill = land.color))
```

Look at them by installations

```{r}
on.base %>% as.data.frame() %>% 
  group_by(instname) %>%
  select(Specimen, GENUS, SPECIES, SDCountyMapID, Year, Month, instname, geometry) %>%
  unite(col = "Sp", GENUS:SPECIES) -> on.base.by.inst

on.base.by.inst %>% group_by(instname, Sp, Year) %>%
  summarise(n = n()) -> on.base.by.inst.summary

ggplot(on.base.by.inst.summary) +
  geom_bar(aes(x = Year, y = n, fill = Sp), stat = "identity") +
  #geom_point(aes(x = Year, y = n, color = Sp)) +
  facet_wrap("instname") +
  scale_fill_viridis_d() +
  theme(axis.title.x = element_blank())

# ggsave("figures/stranding_by_base.png",
#        device = "png", dpi = 600)
```
```{r}
on.base.by.inst.summary %>% data.frame()
```

Just Camp Pendleton:

```{r}
on.base.by.inst %>% filter(instname == "MCB Camp Pendleton") -> on.base.by.inst.CP

on.base.by.inst.CP %>% group_by(instname, Sp, Year) %>%
  summarise(n = n()) -> on.base.by.inst.CP.summary

ggplot(on.base.by.inst.CP.summary) +
  geom_bar(aes(x = Year, y = n, fill = Sp), stat = "identity") +
  #geom_point(aes(x = Year, y = n, color = Sp)) +
  #facet_grid(rows = "instname") +
  scale_fill_viridis_d() +
  ggtitle("Strandings in MCB Camp Pendleton") +
  theme(axis.title.x = element_blank())

```


Every installation along the coast gets some stranding. 

Look at just silver strand

```{r}
ggplot() +
  geom_sf(data = coast.line)+
  geom_sf(data = harbor_bay_1, fill = water.color) +
  geom_sf(data = Pacific, fill = water.color) +
    geom_sf(data = SD.county.line,
          color = "purple",
          fill = "transparent") +

    geom_sf(data = Navy, fill = Navy.color) +

  geom_sf(data = on.base, 
          aes(color = Year), alpha = 0.6) +
  scale_color_viridis_c() +
  coord_sf(xlim = c(W.lon.SS, E.lon.SS),
           ylim = c(S.lat.SS, N.lat.SS)) +

  ylab("Latitude") +
  xlab("Longitude") +
  ggtitle("Cetacean strandings on military installations along Silver Strand (2000-2019)") +
  theme(plot.title = element_text(hjust = 0.5),
        #legend.position = "top", #c(0.1, 0.3),
        legend.background = element_blank(),
        legend.title = element_blank(),
        panel.background = element_rect(fill = land.color))
```


```{r}
on.base.by.inst %>% filter(instname != "MCB Camp Pendleton") -> on.base.by.inst.SS

on.base.by.inst.SS %>% group_by(instname, Sp, Year) %>%
  summarise(n = n()) -> on.base.by.inst.SS.summary

ggplot(on.base.by.inst.SS.summary) +
  geom_bar(aes(x = Year, y = n, fill = Sp), stat = "identity") +
  #geom_point(aes(x = Year, y = n, color = Sp)) +
  facet_grid(rows = "instname") +
  scale_fill_viridis_d() +
  theme(axis.title.x = element_blank())

# ggsave("figures/SS_strandings_by_base.png",
#        device = "png", dpi = 600)
```

Makes sense to look at just Delphins, especially capensis. 


```{r}
dat.1 %>% group_by(GENUS) %>%
  summarise(n = n()) %>% 
  arrange(desc(n)) -> dat.1.bySp

ggplot(data = dat.1.bySp) +
  geom_bar(aes(x = reorder(GENUS, -n), y = n),
                 stat = "identity") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5),
        axis.title.x = element_blank())
```

Needless to say Delphinus and Tursiops are key species to look at. Delphinus has two species.

```{r}
dat.1 %>% filter(GENUS == "Delphinus") -> dat.Delphinus
dat.Delphinus %>%  st_as_sf(coords = c("Longitude", "Latitude"), 
           crs = "+proj=longlat +datum=WGS84") -> dat.Delphinus.sp 

ggplot(dat.Delphinus) +
  geom_histogram(aes(x = Year, fill = SPECIES),
           position = position_dodge()) +
  scale_fill_viridis_d() +
  ggtitle("Delphinus strnadings along SD") +
  theme(axis.title.x = element_blank())

# ggsave("figures/Delphinus_strandings_per_Year.png",
#        device = "png", dpi = 600)
```

```{r}
dat.Delphinus %>% group_by(Year, SPECIES) %>%
  summarize(n = n()) -> Delphinus.byYr

Delphinus.byYr %>% filter(SPECIES  != "sp.") -> Delphinus.byYr.noSp
capensis.vs.delphis <- reshape::cast(Delphinus.byYr.noSp, Year ~ SPECIES, value = "n")

ggplot(capensis.vs.delphis) + 
  geom_point(aes(x = capensis, y = delphis, color = Year)) +
  scale_color_viridis_c()
```

No obvious patterns between the two species. I'll just focus on capensis.


```{r}
dat.Delphinus %>%  filter(SPECIES == "capensis") -> dat.capensis

dat.capensis %>% st_as_sf(coords = c("Longitude", "Latitude"), 
           crs = "+proj=longlat +datum=WGS84") -> dat.capensis.sp 

ggplot() +
  geom_sf(data = coast.line)+
  geom_sf(data = harbor_bay_1, fill = water.color) +
  geom_sf(data = Pacific, fill = water.color) +
    geom_sf(data = SD.county.line,
          color = "purple",
          fill = "transparent") +

    geom_sf(data = Navy, fill = Navy.color) +

  geom_sf(data = dat.capensis.sp, 
          aes(color = Year), alpha = 0.6) +
  scale_color_viridis_c() +
  coord_sf(xlim = c(W.lon, E.lon),
           ylim = c(S.lat, N.lat)) +

  ylab("Latitude") +
  xlab("Longitude") +
  ggtitle("Delphinus capensis strandings along San Diego (2000-2019)") +
  theme(plot.title = element_text(hjust = 0.5),
        #legend.position = "top", #c(0.1, 0.3),
        legend.background = element_blank(),
        legend.title = element_blank(),
        panel.background = element_rect(fill = land.color))

```

```{r}
dat.1.SS %>% filter(SPECIES == "capensis") -> dat.capensis.SS
dat.1.noSS %>% filter(SPECIES == "capensis") -> dat.capensis.noSS

year.df <- data.frame(Year = capensis.vs.delphis$Year)
dat.capensis.SS %>% group_by(Year) %>%
  summarize(n = n(),
            n_per_km = n/(SS.length/1000)) %>%
  right_join(year.df, by = "Year") %>%
  arrange(by = Year) %>%
  mutate(Area = "SS only") -> capensis.SS.byYear

dat.capensis.noSS %>% group_by(Year) %>%
  summarize(n = n(),
            n_per_km = n/((SD.coast.no.MB.SDB.length-SS.length)/1000))  %>%
  right_join(year.df, by = "Year") %>%
  arrange(by = Year) %>%
  mutate(Area = "No SS") -> capensis.noSS.byYear

capensis.vs.delphis %>% 
  select(Year, capensis) %>% 
  transmute("Year" = Year, "n" = capensis, "Area" = "All") %>% #-> tmp
  mutate(n_per_km = n/((SD.coast.no.MB.SDB.length)/1000)) %>%
  rbind(capensis.SS.byYear, capensis.noSS.byYear) -> capensis.all

capensis.all[is.na(capensis.all$n), "n"] <- 0

capensis.vs.delphis %>% 
  select(Year, capensis) %>% 
  transmute("Year" = Year, "n.all" = capensis) %>% #-> tmp
  mutate(n_per_km_all = n.all/((SD.coast.no.MB.SDB.length)/1000)) %>%
  left_join(capensis.SS.byYear, by = "Year") %>%
  transmute("Year" = Year, "n.all" = n.all, "n_per_km_all" = n_per_km_all, 
            "n.SS" = n, "n_per_km_SS" = n_per_km) %>%
  left_join(capensis.noSS.byYear, by = "Year") %>%
  transmute("Year" = Year, "n.all" = n.all, "n_per_km_all" = n_per_km_all,
            "n.SS" = n.SS, "n_per_km_SS" = n_per_km_SS, 
            "n.noSS" = n, "n_per_km_noSS" = n_per_km) -> capensis.all.wide

#write.csv(capensis.all.wide, file = "data/capensis_strandings.csv",
#          row.names = F)

ggplot(capensis.all) +
  geom_bar(aes(x = Year, y = n_per_km, fill = Area), 
           stat = "identity", position = position_dodge()) +
  theme(axis.title.x = element_blank(),
        legend.title = element_blank(),
        legend.position = "top") +
  labs(title = "Stranding of D. capensis") +
  ylab("# strandings per km of coast line") +
  scale_fill_viridis_d()

# ggsave(filename = "figures/capensis_stranding.png",
#        device = "png", dpi = 600)

```

```{r}
dat.capensis.SS %>% st_as_sf(coords = c("Longitude", "Latitude"), 
           crs = "+proj=longlat +datum=WGS84") -> dat.capensis.SS.sp 

ggplot() +
  geom_sf(data = coast.line)+
  geom_sf(data = harbor_bay_1, fill = water.color) +
  geom_sf(data = Pacific, fill = water.color) +
    geom_sf(data = SD.county.line,
          color = "purple",
          fill = "transparent") +

    geom_sf(data = Navy, fill = Navy.color) +

  geom_sf(data = dat.capensis.SS.sp, 
          aes(color = Year), alpha = 0.6) +
  scale_color_viridis_c() +
  coord_sf(xlim = c(W.lon.SS, E.lon.SS),
           ylim = c(S.lat.SS, N.lat.SS)) +

  ylab("Latitude") +
  xlab("Longitude") +
  ggtitle("Delphinus capensis strandings along Silver Strand (2000-2019)") +
  theme(plot.title = element_text(hjust = 0.5),
        #legend.position = "top", #c(0.1, 0.3),
        legend.background = element_blank(),
        legend.title = element_blank(),
        panel.background = element_rect(fill = land.color))
```

Count the number of capensis strandings in and outside of military installations:

```{r}
on.base.by.inst %>% filter(instname != "MCB Camp Pendleton") %>% 
  filter(Sp == "Delphinus_capensis") -> on.base.SS.capensis

dat.capensis.SS.sp %>% st_join(Navy) %>%
  filter(is.na(Shape_Area)) %>% 
  as.data.frame() %>% 
  select(Specimen, GENUS, SPECIES, 
         SDCountyMapID, Year, Month, 
         instname, geometry) %>%
  unite(col = "Sp", GENUS:SPECIES) %>%
  filter(Sp == "Delphinus_capensis") ->  off.base.SS.capensis

# find the coast line length for each base:
Navy %>% filter(instname == "NAS North Island") -> NI.poly
SDB.NI.poly <- st_intersection(SDB.poly, NI.poly)
SDB.NI.line <- st_cast(SDB.NI.poly, "MULTILINESTRING")

NI.line <- st_cast(NI.poly, "MULTILINESTRING")

```



Look at what affects the stranding rate:

NOI: Northern and Southern Oscillation Indices - The NOI (extratropical-based Northern Oscillation Index) and its analog, the SOI* (extratropical-based Southern Oscillation Index) are new indices of midlatitude climate fluctuations that show interesting relationships with fluctuations in marine ecosystems and populations. They reflect the variability in equatorial and extratropical teleconnections and represent a wide range of local and remote climate signals. The indices are counterparts to the traditional SOI (Southern Oscillation Index) that relate variability in the atmospheric forcing of climate change in northern and southern midlatitude hemisphere regions. (https://oceanview.pfeg.noaa.gov/erddap/info/cciea_OC_NOI/index.html)

PDO: Updated standardized values for the PDO index, derived as the leading PC of monthly SST anomalies in the North Pacific Ocean, poleward of 20N. The monthly mean global average SST anomalies are removed to separate this pattern of variability from any "global warming" signal that may be present in the data. (https://oceanview.pfeg.noaa.gov/erddap/info/cciea_OC_PDO/index.html)

```{r}

NOI <- read_csv(file = "data/NOI_v1.csv",
                col_types = cols(time = col_datetime(),
                  NOI = col_double(),
                  SOIX = col_double(),
                  SOI = col_double())) %>%
  mutate(Year = lubridate::year(time))

NOI %>% group_by(Year) %>%
  summarize(NOI = mean(NOI)) -> mean.NOI

PDO <- read_csv(file = "data/PDO_v1.csv",
                col_types = cols(time = col_datetime())) %>%
  mutate(Year = lubridate::year(time))

PDO %>% group_by(Year) %>%
  summarize(PDO = mean(PDO)) -> mean.PDO

capensis.all %>% left_join(mean.NOI, by = "Year") %>%
  left_join(mean.PDO, by = "Year") -> capensis.all.NOI.PDO

ggplot(capensis.all.NOI.PDO) +
  geom_point(aes(x = NOI, y = n_per_km, color = Area))
```

Possibly... there is an increasing relationship with annual averaged NOI and stranding per km of coast line? 

```{r}
ggplot(capensis.all.NOI.PDO) +
  geom_point(aes(x = PDO, y = n_per_km, color = Area))
```

Also, possibly a negative regative relationship with PDO.


```{r}
capensis.all.NOI.PDO %>% filter(Area == "No SS") -> capensis.noSS.NOI.PDO
ggplot(capensis.noSS.NOI.PDO) +
  geom_point(aes(x = NOI, y = n_per_km, color = Year))

```

```{r}
ggplot(capensis.noSS.NOI.PDO) +
  geom_point(aes(x = PDO, y = n_per_km, color = Year))

```


```{r}
fit1 <- lm(n_per_km ~ NOI * PDO, data = capensis.noSS.NOI.PDO)

summary(fit1)
```


```{r}
capensis.all.NOI.PDO %>% filter(Area == "SS only") -> capensis.SS.NOI.PDO
ggplot(capensis.SS.NOI.PDO) +
  geom_point(aes(x = NOI, y = n_per_km, color = Year))

```

```{r}
ggplot(capensis.SS.NOI.PDO) +
  geom_point(aes(x = PDO, y = n_per_km, color = Year))

```


```{r}
fit2 <- lm(n_per_km ~ NOI * PDO, data = capensis.SS.NOI.PDO)

summary(fit2)
```



```{r}
dat.1 %>% filter(GENUS == "Tursiops") -> dat.Tursiops

dat.Tursiops %>% st_as_sf(coords = c("Longitude", "Latitude"), 
           crs = "+proj=longlat +datum=WGS84") -> dat.Tursiops.sp 

ggplot() +
  geom_sf(data = coast.line)+
  geom_sf(data = harbor_bay_1, fill = water.color) +
  geom_sf(data = Pacific, fill = water.color) +
    geom_sf(data = SD.county.line,
          color = "purple",
          fill = "transparent") +

    geom_sf(data = Navy, fill = Navy.color) +

  geom_sf(data = dat.Tursiops.sp, 
          aes(color = Year), alpha = 0.6) +
  scale_color_viridis_c() +
  coord_sf(xlim = c(W.lon, E.lon),
           ylim = c(S.lat, N.lat)) +

  ylab("Latitude") +
  xlab("Longitude") +
  ggtitle("Tursiops strandings along San Diego (2000-2019)") +
  theme(plot.title = element_text(hjust = 0.5),
        #legend.position = "top", #c(0.1, 0.3),
        legend.background = element_blank(),
        legend.title = element_blank(),
        panel.background = element_rect(fill = land.color))

```


```{r}
dat.1.SS %>% filter(GENUS == "Tursiops") -> dat.Tursiops.SS

dat.Tursiops.SS  %>% filter(Latitude < SilverStrandN.lat &
                              Latitude > US.MX.border.lat) %>% 
  st_as_sf(coords = c("Longitude", "Latitude"), 
           crs = "+proj=longlat +datum=WGS84") -> dat.Tursiops.SS.sp 

ggplot() +
  geom_sf(data = coast.line)+
  geom_sf(data = harbor_bay_1, fill = water.color) +
  geom_sf(data = Pacific, fill = water.color) +
    geom_sf(data = SD.county.line,
          color = "purple",
          fill = "transparent") +

    geom_sf(data = Navy, fill = Navy.color) +

  geom_sf(data = dat.Tursiops.SS.sp, 
          aes(color = Year), alpha = 0.6) +
  scale_color_viridis_c() +
  coord_sf(xlim = c(W.lon.SS, E.lon.SS),
           ylim = c(S.lat.SS, N.lat.SS)) +

  ylab("Latitude") +
  xlab("Longitude") +
  ggtitle("Tursiops truncatus strandings along Silver Strand (2000-2019)") +
  theme(plot.title = element_text(hjust = 0.5),
        #legend.position = "top", #c(0.1, 0.3),
        legend.background = element_blank(),
        legend.title = element_blank(),
        panel.background = element_rect(fill = land.color))
```


```{r}
dat.1.noSS %>% filter(GENUS == "Tursiops") -> dat.Tursiops.noSS

dat.Tursiops.SS %>% group_by(Year) %>%
  summarize(n = n(),
            n_per_km = n/(SS.length/1000)) %>%
  right_join(year.df, by = "Year") %>%
  arrange(by = Year) %>%
  mutate(Area = "SS only") -> Tursiops.SS.byYear

dat.Tursiops.noSS %>% group_by(Year) %>%
  summarize(n = n(),
            n_per_km = n/((SD.coast.no.MB.SDB.length-SS.length)/1000))  %>%
  right_join(year.df, by = "Year") %>%
  arrange(by = Year) %>%
  mutate(Area = "No SS") -> Tursiops.noSS.byYear

dat.Tursiops %>% group_by(Year) %>% 
  summarize(n = n()) %>%
  select(Year, n) %>% 
  mutate(n_per_km = n/((SD.coast.no.MB.SDB.length)/1000),
         Area = "All") %>%
  rbind(Tursiops.SS.byYear, Tursiops.noSS.byYear) -> Tursiops.all

Tursiops.all[is.na(Tursiops.all$n), "n"] <- 0

ggplot(Tursiops.all) +
  geom_bar(aes(x = Year, y = n_per_km, fill = Area), 
           stat = "identity", position = position_dodge()) +
  theme(axis.title.x = element_blank(),
        legend.title = element_blank(),
        legend.position = "top") +
  labs(title = "Stranding of Tursiops") +
  ylab("# strandings per km of coast line") +
  scale_fill_viridis_d()

# ggsave(filename = "figures/Tursiops_stranding.png",
#        device = "png", dpi = 600)

```

```{r}
Tursiops.all %>% left_join(mean.NOI, by = "Year") %>%
  left_join(mean.PDO, by = "Year") -> Tursiops.all.NOI.PDO

ggplot(Tursiops.all.NOI.PDO) +
  geom_point(aes(x = NOI, y = n_per_km, color = Area))
```


```{r}
ggplot(Tursiops.all.NOI.PDO) +
  geom_point(aes(x = PDO, y = n_per_km, color = Area))
```

Also, possibly a negative relationship with PDO.


```{r}
Tursiops.all.NOI.PDO %>% filter(Area == "No SS") -> Tursiops.noSS.NOI.PDO
ggplot(Tursiops.noSS.NOI.PDO) +
  geom_point(aes(x = NOI, y = n_per_km, color = Year))

```

```{r}
ggplot(Tursiops.noSS.NOI.PDO) +
  geom_point(aes(x = PDO, y = n_per_km, color = Year))

```


```{r}
fit1 <- lm(n_per_km ~ NOI * PDO, data = Tursiops.noSS.NOI.PDO)

summary(fit1)
```


```{r}
Tursiops.all.NOI.PDO %>% filter(Area == "SS only") -> Tursiops.SS.NOI.PDO
ggplot(Tursiops.SS.NOI.PDO) +
  geom_point(aes(x = NOI, y = n_per_km, color = Year))

```

```{r}
ggplot(Tursiops.SS.NOI.PDO) +
  geom_point(aes(x = PDO, y = n_per_km, color = Year))

```


```{r}
fit2 <- lm(n_per_km ~ NOI * PDO, data = Tursiops.SS.NOI.PDO)

summary(fit2)
```